---
title: "Baltimore Public Schools: Girls of Color"
output: html_document
---

Processing steps for girls of color analysis for Baltimore Public Schools

###Load up denominator of data
```{r}
##Set directory
  setwd("~/Documents/Github/cwg_website/data/")

##Libraries
  library(ggmap)
  library(RColorBrewer)
  library(leaflet)

##Load up data
  denom <- read.delim("sc131a-bcpss.txt")
  inc <- read.csv("incidents_2013.csv")

```


##Denominator: light processing
```{r}
##Denominator from SC131 
  denom <- denom[,c(9,21,25,26,46,47,63,68,72,298:320)]
  denom$SCHNAM <- as.character(denom$SCHNAM)
  denom$id <- 1:nrow(denom) ##Assign ID
  for(i in 10:ncol(denom)){ ##Remove missing values under 0
    denom[,i] <- as.numeric(denom[,i])
    denom[,i][denom[,i]<0]<-NA
  }
  denom$lat <- denom$LATCOD
  denom$lon <- denom$LONCOD

##SChool Level
  denom$LEVEL<-as.character(denom$LEVEL)
  denom$LEVEL[denom$LEVEL=="1"] <- "Primary School"
  denom$LEVEL[denom$LEVEL=="2"] <- "Middle School"
  denom$LEVEL[denom$LEVEL=="3"] <- "High School"
  denom$LEVEL[denom$LEVEL=="4"] <- "Other"

##Grades
  for(i in 5:6){
    denom[,i] <- gsub("0","",as.character(denom[,i]))
    denom[,i][denom[,i]=="PK"]<-"Pre-K"
    denom[,i][denom[,i]=="KG"]<-"K"
  }

##Type
  denom$TYPE <- as.character(denom$TYPE)
  denom$TYPE[denom$TYPE=="1"] <- "Primary School"
  denom$TYPE[denom$TYPE=="2"] <- "Middle School"
  denom$TYPE[denom$TYPE=="3"] <- "High School"
  denom$TYPE[denom$TYPE=="4"] <- "Other"
  
##CHARTER
  denom$CHARTR <- as.character(denom$CHARTR)
  denom$CHARTR[denom$CHARTR=="1"] <-"Yes"
  denom$CHARTR[denom$CHARTR=="2"] <-"No"

##Free lunch
  denom$free_red_lunch<-denom$TOTFRL/denom$MEMBER
```


##Baltimore Incidents: light processing
```{r}
##Incident data from Baltimore Public Schools
  inc <- reshape(inc,
               timevar = "gender",
               idvar = "school",
               direction = "wide")
  inc$SCHNAM <- toupper(inc$school)
  inc$id <- 1:nrow(inc)

```


###Greedy Match Algorithm between Baltimore Public Schools data and SC131a
- Wrote an algorithm to match based on precision of name
- 131 of 132 schools were matched from the BCP data

```{r}
psm <- function(a,b,var,max,min,id){
  
  matched <- data.frame()
  
  for(k in max:min){
    a[,var]<-substr(a[,var],1,k+2)
    b[,var]<-substr(b[,var],1,k+2)
    
    for(j in 1:nrow(a)){
      
      target <- a[j,]
      target <- merge(target,b,by=var)
    
    if(nrow(target)>0){
     # print("if")
      target <- target[1:1,]
      matched<-rbind(matched,cbind(target,k))
      listx <- matched[,paste(id,".x",sep="")]
      listy <- matched[,paste(id,".y",sep="")]
      
      a <- a[ !a[,id] %in% listx, ]
      b <- b[ !b[,id] %in% listy, ]
      
    }
    
    rm(target)
    }
  }
  return(matched)
}
```

131 of 132 schools matched.

```{r}
  result <- psm(inc,denom,"SCHNAM",50,4,"id")
 
```

###Incident rates by race
```{r}
result$pct_aframerf  <- result$inc_aframer.Female/(result$BLALF)
result$pct_amerind  <- result$inc_amerind.Female/result$AMALF
result$pct_aapi  <- result$inc_asian.Female/(result$ASALF+result$HPALF)
result$pct_hisp  <- result$inc_hisp.Female/result$HIALF
result$pct_white <- result$inc_white.Female/result$WHALF

for(i in (ncol(result)-4):ncol(result)){
  print(colnames(result)[i])
  result[,i][result[,i]==Inf]<-NA
  print(summary(result[,i]))
}

```


###Race share of females in school
```{r}
result$total_female <- (result$BLALF +result$AMALF +result$ASALF+result$HPALF +result$HIALF + result$WHALF)
result$share_aframerf  <- result$BLALF/result$total_female
result$share_amerind  <- result$AMALF/result$total_female
result$share_aapi  <- (result$ASALF+result$HPALF )/(result$total_female)
result$share_hisp  <- result$HIALF/result$total_female
result$share_white <- result$WHALF/result$total_female
```


###Incident share of females in school
```{r}
result$totalinc_female <- (result$inc_aframer.Female +result$inc_amerind.Female +result$inc_asian.Female +result$inc_hisp.Female+result$inc_white.Female)
result$shareinc_aframerf  <- result$inc_aframer.Female/result$totalinc_female
result$shareinc_amerind  <- result$inc_amerind.Female/result$totalinc_female
result$shareinc_aapi  <- result$inc_asian.Female/(result$totalinc_female)
result$shareinc_hisp  <- result$inc_hisp.Female/result$totalinc_female
result$shareinc_white <- result$inc_white.Female/result$totalinc_female

```


```{r}

  result <- result[!is.na(result$MEMBER),]

  x_all <-as.numeric(result$lon)
  y_all <- as.numeric(result$lat)
  school <- result$school
  grades <- paste(result$GSLO," - ", result$GSHI,sep="") 
  level <- result$LEVEL
  charter <- result$CHARTR
  free_lunch <- round(100*result$free_red_lunch,1)
  student_body <- result$total_female
  incidents <- result$totalinc_female 

#African Americans
  pr <- round(result$pct_aframerf,1)

  share <- round(100*result$share_aframerf,1)
  shareinc <- round(100*result$shareinc_aframerf,1)
  inc_cnt <- result$inc_aframer.Female
  sb_cnt <- result$BLALF

#Hispanics
  pr_hisp <- round(result$pct_hisp,2)
  share_hisp <- round(100*result$share_hisp,2)
  shareinc_hisp <- round(100*result$shareinc_hisp,1)
  sb_cnt_hisp <- result$HIALF
  inc_cnt_hisp <- result$inc_hisp.Female

#White
  pr_white <- round(result$pct_white,2)
  share_white <- round(100*result$share_white,2)
sb_cnt_white <- result$WHALF
shareinc_white <- round(100*result$shareinc_white,1)
inc_cnt_white <- result$inc_white.Female

#aapi
  pr_asian <- round(result$pct_aapi,2)
  share_asian <- round(100*result$share_aapi,2)
sb_cnt_asian <- result$ASALF+result$HPALF 
shareinc_asian <- round(100*result$shareinc_aapi,1)
inc_cnt_asian <- result$inc_asian.Female


  content_aframer <- paste0("<h3>",school,"</h3>
                            <hr><p><strong>Statistics for Female African American Students </strong><br>
                            <u>Student Body</u>:  <span style='color:red'>  ","<strong> ",sb_cnt,"</strong></span> of <strong>",student_body,"</strong> (",share,"%)<br>
                            <u>Incidents (Total)</u>: <span style='color:red;font-weight:200%'> ","<strong> ",inc_cnt,"</strong></span> of ",incidents," (",shareinc,"%) <br> 
                            <u>Incidents (per pupil)</u>:  <span style='color:red'>  ","<strong> ",pr,"</strong></span><br><br>
                            <strong>School Characteristics </strong><br>
                            <u>Level</u>: ",level,"<br> 
                            <u>Grades</u>: ",grades,"<br> 
                            <u>Charter School</u>: ",charter,"<br> 
                            <u>Free or Reduced Price Lunch</u>: ",free_lunch,"% Eligible<br> 
                            </p>")

 content_hisp <- paste0("<h3>",school,"</h3>
                            <hr><p><strong>Statistics for Female Hispanic Students </strong><br>
                            <u>Student Body</u>:  <span style='color:red'>  ","<strong> ",sb_cnt_hisp,"</strong></span> of <strong>",student_body,"</strong> (",share_hisp,"%)<br>
                            <u>Incidents (Total)</u>: <span style='color:red'> ","<strong> ",inc_cnt_hisp,"</strong></span> of ",incidents," (",shareinc_hisp,"%) <br> 
                            <u>Incidents (per pupil)</u>:  <span style='color:red'>  ","<strong> ",pr_hisp,"</strong></span><br><br>
                            <strong>School Characteristics </strong><br>
                            <u>Level</u>: ",level,"<br> 
                            <u>Grades</u>: ",grades,"<br> 
                            <u>Charter School</u>: ",charter,"<br> 
                            <u>Free or Reduced Price Lunch</u>: ",free_lunch,"% Eligible<br> 
                            </p>") 

 content_white <- paste0("<h3>",school,"</h3>
                            <hr><p><strong>Statistics for Female White Students </strong><br>
                            <u>Student Body</u>:  <span style='color:red'>  ","<strong> ",sb_cnt_white,"</strong></span> of <strong>",student_body,"</strong> (",share_white,"%)<br>
                            <u>Incidents (Total)</u>: <span style='color:red'> ","<strong> ",inc_cnt_white,"</strong></span> of ",incidents," (",shareinc_white,"%) <br> 
                            <u>Incidents (per pupil)</u>:  <span style='color:red'>  ","<strong> ",pr_hisp,"</strong></span><br><br>
                            <strong>School Characteristics </strong><br>
                            <u>Level</u>: ",level,"<br> 
                            <u>Grades</u>: ",grades,"<br> 
                            <u>Charter School</u>: ",charter,"<br> 
                            <u>Free or Reduced Price Lunch</u>: ",free_lunch,"% Eligible<br> 
                            </p>") 


 content_asian <- paste0("<h3>",school,"</h3>
                            <hr><p><strong>Statistics for Female Asian American + Pacific Islanders Students </strong><br>
                            <u>Student Body</u>:  <span style='color:red'>  ","<strong> ",sb_cnt_asian,"</strong></span> of <strong>",student_body,"</strong> (",share_asian,"%)<br>
                            <u>Incidents (Total)</u>: <span style='color:red'> ","<strong> ",inc_cnt_asian,"</strong></span> of ",incidents," (",shareinc_asian,"%) <br> 
                            <u>Incidents (per pupil)</u>:  <span style='color:red'>  ","<strong> ",pr_hisp,"</strong></span><br><br>
                            <strong>School Characteristics </strong><br>
                            <u>Level</u>: ",level,"<br> 
                            <u>Grades</u>: ",grades,"<br> 
                            <u>Charter School</u>: ",charter,"<br> 
                            <u>Free or Reduced Price Lunch</u>: ",free_lunch,"% Eligible<br> 
                            </p>") 


pal <- colorNumeric(
  palette = "Reds",
  domain = pr
)
pal_hisp <- colorNumeric(
  palette = "Blues",
  domain = pr_hisp
)
pal_white<- colorNumeric(
  palette = "Greens",
  domain = pr_white
)
pal_white<- colorNumeric(
  palette = "Greens",
  domain = pr_asian
)
```

###The Maps
To start, we'll initialize a leaflet.js map.

```{r}

leaflet(width="100%") %>% 
  setView(lat = mean(y_all), lng = mean(x_all),12) %>%
  addTiles('http://{s}.basemaps.cartocdn.com/dark_all/{z}/{x}/{y}.png',
             attribution = "Something goes here") %>% 
  addCircleMarkers(data = result, lat = ~ lat, lng = ~ lon,radius=(result$inc_aframer.Female)/4, 
                   fillOpacity = 0.8,stroke = FALSE, 
                   color = ~pal(pr), popup = content_aframer, group="African Americans") %>% 
  addCircleMarkers(data = result, lat = ~ lat, lng = ~ lon,radius=(result$inc_hisp.Female)/4, 
                   fillOpacity = 0.8,stroke = FALSE, 
                   color = ~pal_hisp(pr_hisp), popup = content_hisp, group="Hispanics") %>%
   addCircleMarkers(data = result, lat = ~ lat, lng = ~ lon,radius=(result$inc_white.Female)/4, 
                   fillOpacity = 0.8,stroke = FALSE, 
                   color = ~pal_white(pr_white), popup = content_white, group="Whites") %>% 
   addCircleMarkers(data = result, lat = ~ lat, lng = ~ lon,radius=(result$inc_asian.Female)/4, 
                   fillOpacity = 0.8,stroke = FALSE, 
                   color = ~pal_white(pr_asian), popup = content_asian, group="Asians") %>% 
  addLayersControl(
    overlayGroups = c("African Americans","Hispanics","Whites","Asians"),
    options = layersControlOptions(collapsed = FALSE)
  )
  
```


##Calculate Lorenz Curve

###subset data
```{r}
##White
 lorenz_white <- result[!is.na(result$inc_white.Female) & !is.na(result$pct_white) &result$pct_white>0,c("SCHNAM","inc_white.Female")]
 lorenz_black <- result[!is.na(result$inc_aframer.Female)& !is.na(result$pct_aframer) &result$pct_aframer>0,c("SCHNAM","inc_aframer.Female")]
 lorenz_hisp <- result[!is.na(result$inc_hisp.Female)& !is.na(result$pct_hisp) &  result$pct_hisp>0,c("SCHNAM","inc_hisp.Female")]
 lorenz_aapi <- result[!is.na(result$inc_asian.Female)& !is.na(result$pct_aapi) & result$pct_aapi>0,c("SCHNAM","inc_asian.Female")]
```

###Calculate gini
```{r}
  
lorenz <- function(dataset){
  library(sqldf)
  df<-dataset
  colnames(df)<-c("id","value")
  df<-df[order(df[,2]),]
  df$percentile <-1:nrow(df)
  for(i in seq(0.1,1,0.1)){
    if(i == 0.1){
      df$percentile[df[,2]<=quantile(df[,2],i)]<-10
    }else {
      df$percentile[df[,2]<=quantile(df[,2],i) & df[,2]>quantile(df[,2],i-0.1)]<-i*100
    }
  }

 df_sum <- sqldf("SELECT percentile, SUM(value) value
           FROM df
           GROUP BY percentile")
 df_sum$pct <- df_sum$value/sum(df_sum$value)
 df_sum$cumu <- 0
 for(i in 1:nrow(df_sum)){
   df_sum$cumu[i] <- sum(df_sum[1:i,c("pct")])
 }
 df_sum$neutral <- df_sum$percentile
 df_sum$cumu <- df_sum$cumu*100
 df_sum <- df_sum[,c("percentile","cumu","neutral")]
 df_sum <- rbind(df_sum,data.frame(percentile=0,cumu=0,neutral=0))
 df_sum <- df_sum[order(df_sum$percentile),]
 return(df_sum)
 }


  
white <- lorenz(lorenz_white)[,c("percentile","cumu","neutral")]
black <- lorenz(lorenz_black)[,c("percentile","cumu","neutral")]

plot(white$percentile,white$percentile, type="l", col="black",lty=2 ,
  xlim=c(0, 100), ylim=c(0, 100),xlab = "Cumulative of Schools",ylab = "Cumulative of Incidents")
lines(white$percentile,white$cumu, type="l", col="red")
lines(black$percentile,black$cumu, type="l",col="blue")
legend("topleft", inset=.05, title="Female Students by Race",
    c("white","black"),lty=c(1,1), col= c("red","blue"), horiz=TRUE)

gini <- function(dataset,percentile,cumulative){
    tot <-0
for(i in 2:nrow(dataset)){
  temp<- (dataset[i,c("percentile")]-dataset[i-1,c("percentile")])*(dataset[i,c("cumulative")]+dataset[i-1,c("cumulative")])/2
  tot<-tot+temp
  print(tot)
}
  return(tot/10000)
}
 
asd<-gini(black,"percentile","cumu")

dataset<-black
  tot <-0
for(i in 2:nrow(dataset)){
  temp<- (dataset[i,c("percentile")]-dataset[i-1,c("percentile")])*(dataset[i,c("cumu")]+dataset[i-1,c("cumu")])/2
  tot<-tot+temp
  print(tot)
}
 print(tot/10000)
}

```